Predictive Machine Learning - Exercise Activity
========================================================
The goal of this assignment is to predict the type of exercize performed based on sensor data.  This analysis will use the default values for a random forest model due to its high accuracy and acceptable computation time.  

## Data Exploration
First the data is read in and the training group is broken apart further into training (60%) and testing (40%) groups.

```{r}
#To get reproducable results everytime 
set.seed(85)
library(caret, quietly=TRUE)

#Have r run on multiple cores/threads
library(doParallel, quietly=TRUE)
cl = makeCluster(detectCores())
registerDoParallel(cl)

#Read in the data
training <- read.csv("pml-training.csv")
testing <- read.csv("pml-testing.csv")

#Separate test data into two groups
training <- training[, -c(1)] #First column is redundant
inTrain <- createDataPartition(y = training$classe, p = 0.2, list = FALSE)
train1 <- training[inTrain,]
test <- training[-inTrain,]
```

## Cleaning Data
The next step is to identify features that contain mostly NaNs or will not contribute to the model.  
```{r}
nna <- apply(train1, 2, function(x) {
    1 - sum(is.na(x))/nrow(train1)
}) > 0.03
train1 <- train1[, nna]

#nna <- apply(train1, 2, function(x) {mean(is.na(x))}) < 0.03
#train1 <- train[, nna]

nzv <- nearZeroVar(train1, saveMetrics = TRUE)
exclude_nzv <- rownames(subset(nzv, nzv == TRUE))
train1 <- train1[, -which(names(train1) %in% exclude_nzv)]
```

## Creating the model
```{r}
modFit <- train(classe ~ ., data = train1, method = "rf", importance=TRUE)
```

```{r}


modFit$finalModel$confusion
confusionMatrix(predict(modFit,newdata=test), test$classe)

answers <- predict(modFit, newdata = testing)
#pml_write_files = function(x) {
#  n = length(x)
#  for (i in 1:n) {
#    filename = paste0("./problem_id_", i, ".txt")
#    write.table(x[i], file = filename, quote = FALSE, row.names = FALSE, 
#                col.names = FALSE)
#  }
#}
#pml_write_files(answers)

```


```{r fig.width=9, fig.height=8}
rf.model = modFit$finalModel
varImpPlot(rf.model, n.var = nrow(rf.model$importance[1:25,]), main = "33hi")
```

